Predicting the conformational preferences of proteins using a physics-based free energy method

### 2013-05-08 Paper/paper.tex
upload May5th
May8th

The calculation of protein conformational free energy has been successfully attempted by a number of groups
\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
free energies that does not require knowing a pathway from $A$ to $B$.

@@

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l l}\hline
    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
     T0569       &    2KYW        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
\end{tabular}
\end{center}
\end{table}
\Alberto{Should we comment on T0560? Why is it giving such a high value with such high gdt? It's not
the point to make here, but it is what jumps at me.}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=3.8 in]{T0559.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
        method for target T0559.}
\label{fig:T0559}
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5 in]{T0540.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and two models submitted by different
        prediction groups.}
\label{fig:T0540}
\end{figure}
\Alberto{if the drop down bars/point are a different color it will be easier to see when the crystal
is on top of the axis}
\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}

@@

\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
    $\Delta G{A,B} = \Delta G{A,A^\ast} - \Delta G{B,B^\ast} + \Delta G{A^\ast,B^\ast}$.

@@

One advantage of confinement method is that
none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
to compute. The computational cost for the $56$-residue protein with chameleon sequence that is reported in this article is only $4$ hours 
in a single gpu computer\Alberto{all the different confinements in one GPU?}. All calculations were
performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB
forcefield~\cite{Hornak2006} and the generalized born implicit solvent \cite{Mongan2006}.

@@

\bibitem{Jonsson1998}
Jonsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
Coker, 385 (World Scientific, 1998).

\bibitem{E2007}
E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
barrier-crossing eventsi. J. Chem. Phys. 2007, 126, 164103.

@@

@@

\bibitem{Tyka2006}Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.

>>>>

The calculation of protein conformational free energy has been successfully attempted by a number of groups
\cite{Ytreber2006a,Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
free energies that does not require knowing a pathway from $A$ to $B$.

@@

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l l}\hline
    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
     T0569       &    2KYW        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
\end{tabular}
\end{center}
\end{table}
\Alberto{Should we comment on T0560? Why is it giving such a high value with such high gdt? It's not
the point to make here, but it is what jumps at me.}
\Arijit{I think we have made it clear that crystal structures are often more stable due to salt-bridge effects. Also, we have shown
in other cases that sidechains orientation may play important roles inspite of close GDT-TS. You will see free energy difference between 
two computer generated model is not so high. So, I believe sidechain may play some role here.}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=3.8 in]{T0559.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
        method for target T0559. The backbone region of the predicted structures which differ substantially compare to the
        experimental structure upon superposition are colored brown.}
\label{fig:T0559}
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5 in]{T0540.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and two models submitted by different
        prediction groups.}
\label{fig:T0540}
\end{figure}
\Alberto{if the drop down bars/point are a different color it will be easier to see when the crystal
is on top of the axis}
\Arijit{I will wait for Sarina's figure. I have changed the caption of figure 5 to say why some region are colored brown.}

\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}

@@

\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
    $\Delta G_{A,B} = \Delta G_{A,A^\ast} - \Delta G_{B,B^\ast} + \Delta G_{A^\ast,B^\ast}$.

@@

One advantage of confinement method is that
none of the simulations during the restraining step depends on each other. Therefore, with available computer resources it can be fast 
to compute. The computational cost of a single 20 ns simulation (along the confinement step) of $56$-residue protein 
is only $4$ hours in a single gpu computer  \Alberto{all the different confinements in one GPU?}. All calculations were
performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with the ff99SB
forcefield~\cite{Hornak2006} and the generalized born implicit solvent \cite{Mongan2006}.

@@

\bibitem{Jonsson1998}
J\'{on}sson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
Coker, 385 (World Scientific, 1998).

\bibitem{E2007}
E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
barrier-crossing events. J. Chem. Phys. 2007, 126, 164103.

@@

\bibitem{Ytreber2006a}
Ytreberg, F.M.; Swendsen, R.H.; Zuckerman, D.M. Comparison of free energy methods for molecular systems. J Chem Phys. 2006, 125, 184114.

@@

\bibitem{Tyka2006}
Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.
